---
title: "Fennica Authors in Kanto"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fennica Authors in Kanto}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### Introduction

This document provides an example usage of the search_fennica_authors() function, which links authors from the Fennica catalog to their respective Kanto IDs in Finto's 'Finaf' vocabulary. The function helps enrich metadata for bibliographic records.

```{r message = FALSE, warning = FALSE}
suppressMessages(suppressWarnings({
  if (!requireNamespace("devtools", quietly = TRUE)) {
    install.packages("devtools", quiet = TRUE)
  }
  devtools::install_github("fennicahub/finto", quiet = TRUE)
}))
```

After downloading the package  we select the columns `id`, `Title`, `Author`, `Name_100a`	`Code_1000`	`Code_7000`	`Code_7001`.

```{r message = FALSE, warning = FALSE}
library(finna)
library(finto)
library(dplyr)
library(stringr)
data("fennica_subset",package = "finna")
fennica_subset <- tail(fennica_subset)
knitr::kable(fennica_subset,caption = "few selected fennica data")
```

We used the two main columns we used are `Code_1000` and	`Code_7001` which contain the `asteriID` digits and extract those digits and use them to fetch information from finto/kanto. The columns were renamed as they are originally [MARC 21 format  for bibliographical data](https://marc21.kansalliskirjasto.fi/bib/index.htm) which represent fields `100/0` and `700/o `
So the processed output of the retrieved finto/kanto data can be found using the provided code below.

```{r message = FALSE, warning = FALSE}
authors_df <- finto::process_and_fetch_kanto_info(fennica_subset)
knitr::kable(authors_df, caption = "Authors Data Frame from kanto")
```


### Integrating to fennica 

To integrate the processed KANTO author data with your existing Fennica dataset, you can follow these steps:

1. **Process the KANTO Data**: Use the `process_and_fetch_kanto_info` function to retrieve and process the author information from the KANTO database.

2. **Merge with Fennica Data**: Combine the processed KANTO data with your Fennica dataset based on a common identifier, such as `asteriID`.

Here's how you can achieve this in R:

```{r message = FALSE, warning = FALSE}

# Assuming fennica_subset has columns 'Code_1000' and 'Code_7001'
fennica_subset <- fennica_subset %>%
  mutate(
    asteriID_1000 = str_extract(Code_1000, "\\d{9}"),
    asteriID_7001 = str_extract(Code_7001, "\\d{9}"),
    asteriID = coalesce(asteriID_1000, asteriID_7001)
  ) %>%
  select(-asteriID_1000, -asteriID_7001)  # Remove intermediate columns if not needed

# Merge the processed KANTO data with the Fennica dataset
# Ensure 'asteriID' is the common column for merging
merged_data <- dplyr::left_join(fennica_subset, authors_df, by = "asteriID")

# View the merged dataset
knitr::kable(merged_data, caption = "Merged fennica kanto")

```

**Explanation**:

- **Processing KANTO Data**: The `process_and_fetch_kanto_info` function extracts `asteriID`s from the `fennica_data` and fetches corresponding RDF data from the Finto Skosmos API. It returns a cleaned tibble with the retrieved data, including profession labels.

- **Merging Datasets**: The `dplyr::left_join` function merges the `fennica_data` with the `kanto_authors` data on the `asteriID` column. This ensures that all records from `fennica_data` are retained, and matching author information from `kanto_authors` is added.

**Note**: Ensure that both datasets have the `asteriID` column formatted consistently (e.g., as character strings) to avoid mismatches during the merge.

By following these steps, you can enrich your Fennica dataset with detailed author information from the KANTO database, facilitating more comprehensive analyses and insights. 



